\subsubsection*{Univ.-Prof.\,Dr. Christian Clason}
\label{cv:cclason}

\subsubsection*{Contact details}

University of Graz\\
Department of Mathematics and Scientific Computing\\
Heinrichstrasse 36\\
8010 Graz

\email{c.clason@uni-graz.at}\\
\url{https://imsc.uni-graz.at/clason}\\
\orcid{0000-0002-9948-8426}

\subsubsection*{Academic milestones and positions}

\enlargethispage{1cm}
\begin{tabularx}{\linewidth}{ll}
    Oct. 2001 -- July 2007 & \emph{Research assistant} \\
    & Centre of Mathematical Sciences, \\&Technische Universit채t M체nchen \\
    Dec. 2006 & \emph{Doctoral Degree in Mathematics}, Technische Universit채t M체nchen\\
    July  2007 -- Mar. 2008 & \emph{Postdoctoral research assistant}  \\
    & Department of Mathematics and Scientific Computing, University of Graz\\
    Apr.  2008 -- Sep. 2009 & \emph{University assistant}  \\
    & Department of Mathematics and Scientific Computing, University of Graz\\
    Oct.  2009 -- June 2013 & \emph{Assistant professor} \\
    & Department of Mathematics and Scientific Computing, University of Graz\\
    Dec. 2012 & \emph{Habilitation in Mathematics}, University Graz\\
    June  2013 -- Jan. 2014 & \emph{Associate professor} \\
    & Department of Mathematics and Scientific Computing, University of Graz\\
    Feb.  2014 -- Feb. 2021  & \emph{Professor (W2) of Inverse Problems}\\
    & Faculty of Mathematics, University Duisburg-Essen\\
    Mar.  2021 -- & \emph{Professor of Mathematical Optimization} \\
    & Department of Mathematics and Scientific Computing, University of Graz \\
\end{tabularx}

\subsubsection*{Main areas of research}

\begin{itemize}
    \item{Nonsmooth PDE-constrained optimization}
    \item{Inverse problems, especially parameter identification in PDEs}
    \item{Optimal control of partial differential equations}
    \item{Applications in medicine and medical imaging}
\end{itemize}


\subsubsection*{Ten most important publications}

\newrefsection[cv/clason.bib]

\nocite{cc:COCV:2009a}
\nocite{cc:MAGMA:2010b}
\nocite{cc:AO:2010a}
\nocite{cc:MRM:2011a}
\nocite{cc:SIIMS:2011a}
\nocite{cc:SICON:2012}
\nocite{cc:IP:2015}
\nocite{cc:JMR:2015}
\nocite{cc:MCRF:2017}
\nocite{cc:SIREV:2021}

\begin{refcontext}[sorting=nyt,labelprefix=CC]
    \printbibliography[heading=none]
\end{refcontext}

\endrefsection

\subsubsection*{Additional research achievements}

\begin{enumerate}
    \item 2015: Winner of \emph{ISMRM Challenge on RF Pulse Design (multi-slice)} (with Christoph Aigner, Armin Rund, Karl Kunisch, Rudolf Stollberger)
    \item since Feb. 2014: \emph{Chair} of \www{https://www.ifip.org/bulletin/bulltcs/memtc07.htm}{IFIP Working Group\,7} on Inverse Problems and Imaging
    \item Associate Editor of \emph{Computational Optimization and Applications} (since 2019), \emph{Journal of Mathematical Analysis and Applications} (since 2017), \emph{Journal of Industrial Management Optimization} (since 2015), \emph{Journal of Inverse and Ill-posed Problems} (since 2014)
    \item since Aug. 2019: \emph{Founding and Managing Editor} of diamond open access journal \www{https://jnsao.episciences.org}{Journal on Nonsmooth Analysis and Optimization}
    \item since Mar. 2021: \emph{Speaker of consortium} \enquote{Optimization and Numerical Analysis}, Doctoral Academy Graz
    \item since September 2024: \emph{Chair} of \www{https://www.ifip.org/bulletin/bulltcs/memtc07.htm}{IFIP Technical Committee 7} on System Modeling and Optimization
    \item Coordinator of Special Research Area (SFB) \www{https://imsc.uni-graz.at/mr-dynamo/}``Mathematics of Reconstruction in Dynamical and Active Models'' Austrian Science Fund (FWF) grant \www{https://dx.doi.org/10.55776/F100800}{10.55776/F100800}, 2025-2029
    \item Supervision of four PhD students (completed)
\end{enumerate}

